---
title: "1.5.3"
output: html_notebook
---

# Problem 1.1 - Loading and Summarizing the Dataset
How many interviewers re in the dataset?
```{r}
CPS = read.csv("data/CPSData.csv")
summary(CPS)
```
There are 131302 interviewrs

# Problem 1.2 - Loading and Summarizing the Dataset
Among the interviewees with a value reported for the Industry variable, what is the most common industry of employment? Please enter the name exactly how you see it.
```{r}
sort(table(CPS$Industry))
```
Educational and health services

# Problem 1.3 - Loading and Summarizing the Dataset
Recall from the homework assignment "The Analytical Detective" that you can call the sort() function on the output of the table() function to obtain a sorted breakdown of a variable. For instance, sort(table(CPS$Region)) sorts the regions by the number of interviewees from that region.

Which state has the fewest interviewees?
```{r}
sort(table(CPS$State))
```
New Mexico has the fewest interviewers while California the largest

# Problem 1.4 - Loading and Summarizing the Dataset
What proportion of interviewees are citizens of the United States?
```{r}
table(CPS$Citizenship)
```
```{r}
116639 + 7073 + 7590
```
```{r}
(116639 + 7073)/131302
```

# Problem 1.5 - Loading and Summarizing the Dataset
The CPS differentiates between race (with possible values American Indian, Asian, Black, Pacific Islander, White, or Multiracial) and ethnicity. A number of interviewees are of Hispanic ethnicity, as captured by the Hispanic variable. For which races are there at least 250 interviewees in the CPS dataset of Hispanic ethnicity? (Select all that apply.)
```{r}
table(CPS$Race, CPS$Hispanic)
```

# Problem 2.1 - Evaluating Missing Values
Which variables have at least one interviewee with a missing (NA) value? (Select all that apply.)
```{r}
summary(CPS)
```


```{r}
sum(is.na(CPS$Region))
sum(is.na(CPS$State))
sum(is.na(CPS$Married))
sum(is.na(CPS$Sex))
sum(is.na(CPS$Education))
sum(is.na(CPS$Race))
sum(is.na(CPS$Citizenship))
sum(is.na(CPS$EmploymentStatus))
sum(is.na(CPS$Industry))    
    
```

# Problem 2.2 - Evaluating Missing Values
Often when evaluating a new dataset, we try to identify if there is a pattern in the missing values in the dataset. We will try to determine if there is a pattern in the missing values of the Married variable. The function is.na(CPS$Married) returns a vector of TRUE/FALSE values for whether the Married variable is missing. We can see the breakdown of whether Married is missing based on the reported value of the Region variable with the function table(CPS$Region, is.na(CPS$Married)).
```{r}
table(CPS$Region, is.na(CPS$Married))
table(CPS$Sex, is.na(CPS$Married))
table(CPS$Citizenship, is.na(CPS$Married))
table(CPS$Age, is.na(CPS$Married))
```

# Problem 2.3 - Evaluating Missing Values
As mentioned in the variable descriptions, MetroAreaCode is missing if an interviewee does not live in a metropolitan area. Using the same technique as in the previous question, answer the following questions about people who live in non-metropolitan areas.

How many states had all interviewees living in a non-metropolitan area (aka they have a missing MetroAreaCode value)? For this question, treat the District of Columbia as a state (even though it is not technically a state).
```{r}
table(CPS$State, is.na(CPS$MetroAreaCode))
```

# Problem 2.4 - Evaluating Missing Values
Which region of the United States has the largest proportion of interviewees living in a non-metropolitan area?
```{r}
prop.table(table(CPS$Region, is.na(CPS$MetroAreaCode)))
```

# Problem 2.5 - Evaluating Missing Values
While we were able to use the table() command to compute the proportion of interviewees from each region not living in a metropolitan area, it was somewhat tedious (it involved manually computing the proportion for each region) and isn't something you would want to do if there were a larger number of options. It turns out there is a less tedious way to compute the proportion of values that are TRUE. The mean() function, which takes the average of the values passed to it, will treat TRUE as 1 and FALSE as 0, meaning it returns the proportion of values that are true. For instance, mean(c(TRUE, FALSE, TRUE, TRUE)) returns 0.75. Knowing this, use tapply() with the mean function to answer the following questions:

- Which state has a proportion of interviewees living in a non-metropolitan area closest to 30%?
```{r}
sort(tapply(is.na(CPS$MetroAreaCode), CPS$State, mean))
```
Wisconsin

- Which state has the largest proportion of non-metropolitan interviewees, ignoring states where all interviewees were non-metropolitan?
```{r}
sort(tapply(is.na(CPS$MetroAreaCode), CPS$State, mean))
```
Montana

# Problem 3.1 - Integrating Metropolitan Area Data
Codes like MetroAreaCode and CountryOfBirthCode are a compact way to encode factor variables with text as their possible values, and they are therefore quite common in survey datasets. In fact, all but one of the variables in this dataset were actually stored by a numeric code in the original CPS datafile.

When analyzing a variable stored by a numeric code, we will often want to convert it into the values the codes represent. To do this, we will use a dictionary, which maps the the code to the actual value of the variable. We have provided dictionaries MetroAreaCodes.csv and CountryCodes.csv, which respectively map MetroAreaCode and CountryOfBirthCode into their true values. Read these two dictionaries into data frames MetroAreaMap and CountryMap.
```{r}
MetroAreaMap = read.csv("data/MetroAreaCodes.csv")
CountryMap = read.csv("data/CountryCodes.csv")
```
- How many observations (codes for metropolitan areas) are there in MetroAreaMap?
```{r}
nrow(MetroAreaMap)
```
There are 8661480 observation

- How many observations (codes for countries) are there in CountryMap?
```{r}
nrow(CountryMap)
```

# Problem 3.2 - Integrating Metropolitan Area Data
To merge in the metropolitan areas, we want to connect the field MetroAreaCode from the CPS data frame with the field Code in MetroAreaMap. The following command merges the two data frames on these columns, overwriting the CPS data frame with the result:
```{r}
CPS = merge(CPS, MetroAreaMap, by.x="MetroAreaCode", by.y="Code", all.x=TRUE)
```
 What is the name of the variable that was added to the data frame by the merge() operation?
```{r}
summary(CPS)
```

How many interviewees have a missing value for the new metropolitan area variable? Note that all of these interviewees would have been removed from the merged data frame if we did not include the all.x=TRUE parameter.
```{r}
sum(is.na(CPS$MetroArea))
```
 
 # Problem 3.3 - Integrating Metropolitan Area Data
Which of the following metropolitan areas has the largest number of interviewees?
```{r}
sort(table(CPS$MetroArea))
```

# Problem 3.4 - Integrating Metropolitan Area Data
Which metropolitan area has the highest proportion of interviewees of Hispanic ethnicity? Hint: Use tapply() with mean, as in the previous subproblem. Calling sort() on the output of tapply() could also be helpful here.
```{r}
sort(tapply(CPS$Hispanic, CPS$MetroArea, mean))
```

# Problem 3.5 - Integrating Metropolitan Area Data
Remembering that CPS$Race == "Asian" returns a TRUE/FALSE vector of whether an interviewee is Asian, determine the number of metropolitan areas in the United States from which at least 20% of interviewees are Asian.
```{r}
sort(tapply(CPS$Race == "Asian", CPS$MetroArea, mean))
```

# Problem 3.6 - Integrating Metropolitan Area Data
Normally, we would look at the sorted proportion of interviewees from each metropolitan area who have not received a high school diploma with the command:

sort(tapply(CPS$Education == "No high school diploma", CPS$MetroArea, mean))

However, none of the interviewees aged 14 and younger have an education value reported, so the mean value is reported as NA for each metropolitan area. To get mean (and related functions, like sum) to ignore missing values, you can pass the parameter na.rm=TRUE. Passing na.rm=TRUE to the tapply function, determine which metropolitan area has the smallest proportion of interviewees who have received no high school diploma
```{r}
sort(tapply(CPS$Education == "No high school diploma", CPS$MetroArea, mean, na.rm=TRUE))
```

# Problem 4.1 - Integrating Country of Birth Data
Just as we did with the metropolitan area information, merge in the country of birth information from the CountryMap data frame, replacing the CPS data frame with the result. If you accidentally overwrite CPS with the wrong values, remember that you can restore it by re-loading the data frame from CPSData.csv and then merging in the metropolitan area information using the command provided in the previous subproblem.

What is the name of the variable added to the CPS data frame by this merge operation?

```{r}
CPS = merge(CPS, CountryMap, by.x="CountryOfBirthCode", by.y="Code", all.x=TRUE)
summary(CPS)
```
```{r}
sum(is.na(CPS$Country))
```

# Problem 4.2 - Integrating Country of Birth Data
Among all interviewees born outside of North America, which country was the most common place of birth?

```{r}
sort(table(CPS$Country))
```

# Problem 4.3 - Integrating Country of Birth Data
What proportion of the interviewees from the "New York-Northern New Jersey-Long Island, NY-NJ-PA" metropolitan area have a country of birth that is not the United States? For this computation, don't include people from this metropolitan area who have a missing country of birth.
```{r}
table(CPS$MetroArea == "New York-Northern New Jersey-Long Island, NY-NJ-PA", CPS$Country != "United States")
```
```{r}
1668/(1668+3736)
```

# Problem 4.4 - Integrating Country of Birth Data
Which metropolitan area has the largest number (note -- not proportion) of interviewees with a country of birth in India? Hint -- remember to include na.rm=TRUE if you are using tapply() to answer this question.

```{r}
sort(tapply(CPS$Country == "India", CPS$MetroArea, sum, na.rm=TRUE))

sort(tapply(CPS$Country == "Brazil", CPS$MetroArea, sum, na.rm=TRUE))

sort(tapply(CPS$Country == "Somalia", CPS$MetroArea, sum, na.rm=TRUE))
```

