---
title: "Analytics Edge: Unit 1 - Stocks Dynamics"
author: "Antonella Schiavoni"
output: html_notebook
---

# Download and read the following files into R, using the read.csv function
```{r}
IBM = read.csv("IBMStock.csv")
GE = read.csv("GEStock.csv")
ProcterGamble = read.csv("ProcterGambleStock.csv")
CocaCola = read.csv("CocaColaStock.csv")
Boeing = read.csv("BoeingStock.csv")
```

# Problem 1.1 - Summary Statistics
We need to convert the dates into a format that R can understand. Right now, the date variable is stored as a factor. We can convert this to a "Date" object in R.
```{r}
IBM$Date = as.Date(IBM$Date, "%m/%d/%y")

GE$Date = as.Date(GE$Date, "%m/%d/%y")

CocaCola$Date = as.Date(CocaCola$Date, "%m/%d/%y")

ProcterGamble$Date = as.Date(ProcterGamble$Date, "%m/%d/%y")

Boeing$Date = as.Date(Boeing$Date, "%m/%d/%y")
```

# Problem 1.2 - Summary Statistics
What is the earliest year in our datasets?
```{r}
min(IBM$Date)
min(GE$Date)
min(CocaCola$Date)
min(ProcterGamble$Date)
min(Boeing$Date)
```
# Problem 1.3 - Summary Statistics
What is the latests year in our datasets?
```{r}
max(IBM$Date)
max(GE$Date)
max(CocaCola$Date)
max(ProcterGamble$Date)
max(Boeing$Date)
```

# Problem 1.4 - Summary Statistics
What is the mean stock price of IBM over this time period?

```{r}
mean(IBM$StockPrice)
```

# Problem 1.5 - Summary Statistics
What is the minimum stock price of General Electric (GE) over this time period?
```{r}
min(GE$StockPrice)
```

# Problem 1.6 - Summary Statistics
What is the maximum stock price of Coca-Cola over this time period?
```{r}
max(CocaCola$StockPrice)
```

# Problem 1.7 - Summary Statistics
What is the median stock price of Boeing over this time period?
```{r}
median(Boeing$StockPrice)
```

# Problem 1.8 - Summary Statistics
What is the standard deviation of the stock price of Procter & Gamble over this time period?
```{r}
sd(ProcterGamble$StockPrice)
```

# Problem 2.1 - Visualizing Stock Dynamics
Let's plot the stock prices to see if we can visualize trends in stock prices during this time period. Using the plot function, plot the Date on the x-axis and the StockPrice on the y-axis, for Coca-Cola.

Around what year did Coca-Cola has its highest stock price in this time period?
```{r}
plot(CocaCola$Date, CocaCola$StockPrice, type='l')
```
Around 1973 CocaCola shows the highest stock price.
Around 1980 CocaCola shows the lowest stock price

# Problem 2.2 - Visualizing Stock Dynamics
Now, let's add the line for Procter & Gamble too. You can add a line to a plot in R by using the lines function instead of the plot function.

In March of 2000, the technology bubble burst, and a stock market crash occurred. According to this plot, which company's stock dropped more?
```{r}
plot(CocaCola$Date, CocaCola$StockPrice, type='l', col='red')
lines(ProcterGamble$Date, ProcterGamble$StockPrice, col='blue')
abline(v=as.Date(c("2000-03-01")), lwd=2)
```
Both lines drop, but Procter and Gamble drops more.

# Problem 2.3 - Visualizing Stock Dynamics

Around 1983, the stock for one of these companies (Coca-Cola or Procter and Gamble) was going up, while the other was going down. Which one was going up? CocaCola

In the time period shown in the plot, which stock generally has lower values? CocaCola

# Problem 3.1 - Visualizing Stock Dynamics 1995-2005
Which stock fell the most right after the technology bubble burst in March 2000?
```{r}
plot(CocaCola$Date[301:432], CocaCola$StockPrice[301:432], type="l", col="red", ylim=c(0,210), xlab = "Date", ylab = "Stock Price")
lines(ProcterGamble$Date[301:432], ProcterGamble$StockPrice[301:432], col="blue")
lines(IBM$Date[301:432], IBM$StockPrice[301:432], col="green")
lines(GE$Date[301:432], GE$StockPrice[301:432], col="purple")
lines(Boeing$Date[301:432], Boeing$StockPrice[301:432], col="orange")
```
General Electric

# Problem 3.2 - Visualizing Stock Dynamics 1995-2005
Which stock reaches the highest value in the time period 1995-2005? 
IBM

# Problem 3.3 - Visualizing Stock Dynamics 1995-2005
In October of 1997, there was a global stock market crash that was caused by an economic crisis in Asia. Comparing September 1997 to November 1997, which companies saw a decreasing trend in their stock price? (Select all that apply.)

```{r}
plot(CocaCola$Date[301:432], CocaCola$StockPrice[301:432], type="l", col="red", ylim=c(0,210), xlab = "Date", ylab = "Stock Price")
lines(ProcterGamble$Date[301:432], ProcterGamble$StockPrice[301:432], col="blue")
lines(IBM$Date[301:432], IBM$StockPrice[301:432], col="green")
lines(GE$Date[301:432], GE$StockPrice[301:432], col="purple")
lines(Boeing$Date[301:432], Boeing$StockPrice[301:432], col="orange")
abline(v=as.Date(c("1997-09-01")), lwd=2)
abline(v=as.Date(c("1997-11-01")), lwd=2)
```

# Problem 3.4 - Visualizing Stock Dynamics 1995-2005
In the last two years of this time period (2004 and 2005) which stock seems to be performing the best, in terms of increasing stock price?
```{r}
plot(CocaCola$Date[301:432], CocaCola$StockPrice[301:432], type="l", col="red", ylim=c(0,210), xlab = "Date", ylab = "Stock Price")
lines(ProcterGamble$Date[301:432], ProcterGamble$StockPrice[301:432], col="blue")
lines(IBM$Date[301:432], IBM$StockPrice[301:432], col="green")
lines(GE$Date[301:432], GE$StockPrice[301:432], col="purple")
lines(Boeing$Date[301:432], Boeing$StockPrice[301:432], col="orange")
abline(v=as.Date(c("2004-01-01")), lwd=2)
abline(v=as.Date(c("2005-01-01")), lwd=2)
```
Boeing

# Problem 4.1 - Monthly Trends
For IBM, compare the monthly averages to the overall average stock price. In which months has IBM historically had a higher stock price (on average)? Select all that apply.
```{r}
mean(IBM$StockPrice)
tapply(IBM$StockPrice, months(IBM$Date), mean)
```

# Problem 4.2 - Monthly Trends
```{r}
mean(CocaCola$StockPrice)
tapply(CocaCola$StockPrice, months(CocaCola$Date), mean)
```

# Problem 4.3 - Monthly Trends
For the months of December and January, every company's average stock is higher in one month and lower in the other. In which month are the stock prices lower?
```{r}
tapply(GE$StockPrice, months(GE$Date), mean)
tapply(CocaCola$StockPrice, months(CocaCola$Date), mean)
tapply(IBM$StockPrice, months(IBM$Date), mean)
tapply(ProcterGamble$StockPrice, months(ProcterGamble$Date), mean)
tapply(Boeing$StockPrice, months(Boeing$Date), mean)
```
