---
title: "An Analytical Detective"
output: html_notebook
---
List of descriptions of the variables:

ID: a unique identifier for each observation
Date: the date the crime occurred
LocationDescription: the location where the crime occurred
Arrest: whether or not an arrest was made for the crime (TRUE if an arrest was made, and FALSE if an arrest was not made)
Domestic: whether or not the crime was a domestic crime, meaning that it was committed against a family member (TRUE if it was domestic, and FALSE if it was not domestic)
Beat: the area, or "beat" in which the crime occurred. This is the smallest regional division defined by the Chicago police department.
District: the police district in which the crime occured. Each district is composed of many beats, and are defined by the Chicago Police Department.
CommunityArea: the community area in which the crime occurred. Since the 1920s, Chicago has been divided into what are called "community areas", of which there are now 77. The community areas were devised in an attempt to create socially homogeneous regions.
Year: the year in which the crime occurred.
Latitude: the latitude of the location at which the crime occurred.
Longitude: the longitude of the location at which the crime occurred.


Problem 1.1 - Loading the Data
Read the dataset mvtWeek1.csv into R, using the read.csv function, and call the data frame "mvt". Remember to navigate to the directory on your computer containing the file mvtWeek1.csv first. It may take a few minutes to read in the data, since it is pretty large. Then, use the str and summary functions to answer the following questions.

How many rows of data (observations) are in this dataset?
```{r}
mvt <- read.csv("data/analytical-detective/mvtWeek1.csv")
str(mvt)
```
There are 191641 rows of data in the dataset


Problem 1.2 - Loading the Data
How many variables are in this dataset?

There are 11 variables

Problem 1.3 - Loading the Data
Using the "max" function, what is the maximum value of the variable "ID"?
```{r}
max(mvt$ID)
```
The maximum value of ID column is 9181151


Problem 1.4 - Loading the Data
What is the minimum value of the variable "Beat"?
```{r}
min(mvt$Beat)
```
The minimum value of Beat column is 111


Problem 1.5 - Loading the Data
How many observations have value TRUE in the Arrest variable (this is the number of crimes for which an arrest was made)?
```{r}
sum(mvt$Arrest == TRUE)
```
The number of crimes for which an arrest was made is 15536


Problem 1.6 - Loading the Data
How many observations have a LocationDescription value of ALLEY?
```{r}
sum(mvt$LocationDescription == "ALLEY")
```

Problem 2.1 - Understanding Dates in R
In many datasets, like this one, you have a date field. Unfortunately, R does not automatically recognize entries that look like dates. We need to use a function in R to extract the date and time. Take a look at the first entry of Date (remember to use square brackets when looking at a certain entry of a variable).

In what format are the entries in the variable Date?
```{r}
mvt$Date[1]
```
The data format is in Month/Day/Year Hour:Minute



Problem 2.2 - Understanding Dates in R
Now, let's convert these characters into a Date object in R. In your R console, type

DateConvert = as.Date(strptime(mvt$Date, "%m/%d/%y %H:%M"))

This converts the variable "Date" into a Date object in R. Take a look at the variable DateConvert using the summary function.

What is the month and year of the median date in our dataset? Enter your answer as "Month Year", without the quotes. (Ex: if the answer was 2008-03-28, you would give the answer "March 2008", without the quotes.)

```{r}
DateConvert = as.Date(strptime(mvt$Date, "%m/%d/%y %H:%M"))

summary(DateConvert)
```
The median is May 2006


Problem 2.3 - Understanding Dates in R
Now, let's extract the month and the day of the week, and add these variables to our data frame mvt. We can do this with two simple functions. Type the following commands in R:

mvt$Month = months(DateConvert)

mvt$Weekday = weekdays(DateConvert)

This creates two new variables in our data frame, Month and Weekday, and sets them equal to the month and weekday values that we can extract from the Date object. Lastly, replace the old Date variable with DateConvert by typing:

mvt$Date = DateConvert

Using the table command, answer the following questions.

In which month did the fewest motor vehicle thefts occur?
```{r}
mvt$Month = months(DateConvert)
mvt$Weekday = weekdays(DateConvert)
mvt$Date = DateConvert

table(mvt$Month)
```
In February


Problem 2.4 - Understanding Dates in R
On which weekday did the most motor vehicle thefts occur?
```{r}
table(mvt$Weekday)
```
In Friday


Problem 2.5 - Understanding Dates in R
Each observation in the dataset represents a motor vehicle theft, and the Arrest variable indicates whether an arrest was later made for this theft. Which month has the largest number of motor vehicle thefts for which an arrest was made?
```{r}
table(mvt$Month, mvt$Arrest)
```
January is the month has the largest number of motor vehicle thefts for which an arrest was made


Problem 3.1 - Visualizing Crime Trends
Now, let's make some plots to help us better understand how crime has changed over time in Chicago. Throughout this problem, and in general, you can save your plot to a file. For more information, this website very clearly explains the process.

First, let's make a histogram of the variable Date. We'll add an extra argument, to specify the number of bars we want in our histogram. In your R console, type

hist(mvt$Date, breaks=100)

Looking at the histogram, answer the following questions.

In general, does it look like crime increases or decreases from 2002 - 2012?

```{r}
hist(mvt$Date, breaks=100)
```
In general aspects, it looks like it's decreasing

In general, does it look like crime increases or decreases from 2005 - 2008?
It decreases

In general, does it look like crime increases or decreases from 2009 - 2011?
It increases


Problem 3.2 - Visualizing Crime Trends
Now, let's see how arrests have changed over time. Create a boxplot of the variable "Date", sorted by the variable "Arrest" (if you are not familiar with boxplots and would like to learn more, check out this tutorial). In a boxplot, the bold horizontal line is the median value of the data, the box shows the range of values between the first quartile and third quartile, and the whiskers (the dotted lines extending outside the box) show the minimum and maximum values, excluding any outliers (which are plotted as circles). Outliers are defined by first computing the difference between the first and third quartile values, or the height of the box. This number is called the Inter-Quartile Range (IQR). Any point that is greater than the third quartile plus the IQR or less than the first quartile minus the IQR is considered an outlier.

Does it look like there were more crimes for which arrests were made in the first half of the time period or the second half of the time period? (Note that the time period is from 2001 to 2012, so the middle of the time period is the beginning of 2007.)

```{r}
boxplot(mvt$Date ~ mvt$Arrest)
```
In the first half. This is due to the fact that the TRUE boxplot, is skewed towards the bottom of the plot this means that more crimes for which arrest were made in the first half of the time period.


Problem 3.3 - Visualizing Crime Trends
Let's investigate this further. Use the table function for the next few questions.

For what proportion of motor vehicle thefts in 2001 was an arrest made?

Note: in this question and many others in the course, we are asking for an answer as a proportion. Therefore, your answer should take a value between 0 and 1.
```{r}
table(mvt$Arrest, mvt$Year)
```
```{r}
2152/(2152 + 18517)
```

Problem 3.4 - Visualizing Crime Trends
For what proportion of motor vehicle thefts in 2007 was an arrest made?
```{r}
1212/(1212+13068)
```

Problem 3.5 - Visualizing Crime Trends
1/1 punto (calificado)
For what proportion of motor vehicle thefts in 2012 was an arrest made?
```{r}
550/(550 + 13542)
```

